

\documentclass[a4paper]{article}

\usepackage{amsmath}

\newcommand{\expval}[1]{\langle #1 \rangle}
\DeclareMathOperator{\Err}{Err}
\DeclareMathOperator{\erf}{erf}

\title{Analytical results for tests}

\begin{document}

\maketitle

\begin{abstract}
We summarise the analytical results used in the tests for
\verb=scikit-monaco=. 
\end{abstract}

\section{Uniform Sampling MC}

Following Numerical Recipes, we calculate the standard error in the Monte-Carlo
integration of the function $f$ as:

\begin{equation}
    \Err_N(f) = \Omega \sqrt{\frac{\expval{f^2}-\expval{f}^2}{N}}
\end{equation}

where $N$ is the number of points and $\expval{g} = \int_\Omega g(x) dx$, where
$\Omega$ is the volume being sampled during the integration and $x$ denotes all
the variables of integration.


\subsection{Constant function}

Let $f(x) = 1$. Then, $\expval{f} = \Omega$, where $\Omega$ is the volume of
integration. $\expval{f^2} = 1$, such that $\Err_N = 0$ for all $N > 0$.


\subsection{Product function}

Let $f(x) = \prod_i^d x_i$, where $d$ is the dimensionality of the
integration. Thus, if $d = 2$, $f(x,y) = xy$. We consider the $d$-dimensional
hypercube with upper and lower limit $b$ and $a$, respectively, such that each
$a \le x_i \le b$. 

\begin{eqnarray}
    \expval{f} &=& \idotsint_a^b \prod_i x_i dx_i =
    \left(\frac{b^2-a^2}{2}\right)^d \\
    \expval{f^2} &=& \idotsint_a^b \prod_i x_i^2 dx_i =
    \left(\frac{b^3-a^3}{3}\right)^d 
\end{eqnarray}

If $a = 0$ and $b=1$, we have $\expval{f} = 1/2^d$ and $\expval{f^2} = 1/3^d$.
Then, $\Err_N(f) = \frac{\sqrt{1/3^d - 1/4^d}}{N}$.


\subsection{Gaussian}

Let $f(x) = \prod_i^d \exp(-\beta^2 x_i^2)$. Again, we consider the
hypercube such that $a \le x_i \le b$ for each $i$. Then:

\begin{eqnarray}
    \expval{f} &=& \frac{\sqrt{\pi}}{2\beta} (\erf(\beta b) - \erf(\beta a) \\
    \expval{f^2} &=& \frac{\sqrt{\pi/2}}{2\beta} (\erf(\sqrt{2}\beta b) -
    \erf(\sqrt{2}\beta a)
\end{eqnarray}


\section{MISER Monte-Carlo}

Testing that the MISER algorithm returns the correct value (within error bars)
is trivial, provided the integral can be calculated by some ``trusted'' means.

Testing that the MISER algorithm returns the correct error is more difficult.
Unlike for simple uniform sampling case, there is no easy formula for the
calculation of the error. What we can do, though, is make sure that the error
predicted is correct, \emph{on average}. 

Let's assume that we are estimating an integral $I$. Let $I^*$ denote the
correct, analytical value of $I$. 

Pre test

$\langle I \rangle$ is the average of all the MC calls.

$\expval{\sigma}$ is the average of the errors.

$S_I$ is the standard error in $I-I^*$.

\begin{eqnarray}
    \langle I \rangle &=& I^* \pm \langle\sigma\rangle/\sqrt{N} \\
    \langle \sigma \rangle &=& S_I \pm S_I / \sqrt{2(N-1)}
\end{eqnarray}

Then, if those tests pass, extract:

$\expval{\sigma}$ : allowable error.

$\expval{S_\sigma}$ : allowable error in error.


\end{document}
